---
title: "Example code for ggplot2"
author: "Kevin Rue-Albrecht"
date: "03/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggrepel)
library(cowplot)
library(patchwork)
library(gridExtra)
```

# Exercise

Use `library(ggplot2)` to load the package.

```{r}
library(ggplot2)
```

## Specify the dataset.

What does `ggplot(diamonds)` do?

```{r}
ggplot(diamonds)
```

> Answer:

## Add the aesthetics.

What does `ggplot(diamonds, aes(x = carat, y = price))` do?

```{r}
ggplot(diamonds, aes(x = carat, y = price))
```

> Answer:

## Add geometric objects

- Add data points showing `carat` on the x-axis and `price` on the y-axis.

```{r}
ggplot(diamonds, aes(x = carat, y = price))+
  geom_point()
  
```

- Color data points by `cut`.
  (Copy-paste and extend the code chunk above.)

```{r}
ggplot(diamonds, aes(x = carat, y = price, color = cut))+
  geom_point()
  
```

- Add a smoothed mean trend line.
  (Copy-paste and extend the code chunk above.)

```{r}
ggplot(diamonds, aes(x = carat, y = price, color = cut))+
  geom_point()+
  geom_smooth()


```

```{r}
ggplot(diamonds, aes(x = carat, y = price))+
  geom_point(aes(color = cut))+
  geom_smooth()
```


- Assign that last plot to an object called `obds_diamonds`.

```{r}
obds_diamonds <- ggplot(diamonds, aes(x = carat, y = price, color = cut))+
  geom_point()+
  geom_smooth()


```


# Exercise

Predict the difference between these two plots

### Plot 1

```{r, include=TRUE}
ggplot(diamonds, aes(x = carat, y = price, colour = cut)) +
  geom_point() +
  geom_smooth()
```

### Plot 2

```{r, include=TRUE}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point(aes(colour = cut)) +
  geom_smooth()
```

> Answer:

# Exercise

## Trend lines

Using the `ChickWeight` data set:

- Create a scatter plot of weight (y-axis) over time (x-axis).

```{r}
head(ChickWeight)
ggplot( ChickWeight, aes(x = Time, y = weight, color = Diet) )+
  geom_point()+
  geom_smooth()

```

- Color by diet.
  (Copy-paste and extend the code chunk above.)

```{r}
ggplot( ChickWeight, aes(x = Time, y = weight, color = Diet) )+
  geom_point()+
  geom_smooth(method = lm) #method for smooth is linear

```

- Add a linear mean trend line for each diet.
  (Copy-paste and extend the code chunk above.)

```{r}



```

At this point you should be able to visually identify
the diet that leads to the highest mean increase in weight.

> Answer:

- Facet a ribbon of sub-plots, one per diet.
  (Copy-paste and extend the code chunk above.)

```{r}

ggplot( ChickWeight, aes(x = Time, y = weight, color = Diet) )+
  geom_point()+
  geom_smooth(method = lm)+
  facet_wrap(~Diet)


```
```{r}
ggplot( ChickWeight, aes(x = Time, y = weight, color = Diet) )+
  geom_bar()+

```


- Assign that last plot to an object called `obds_chickweight`.

```{r}
obds_chickweight <- ggplot(ChickWeight, aes(x = Time, y = weight, color = Diet))+
  geom_point()+
  geom_smooth(method = lm)+
  facet_wrap(~Diet)



```

# Exercise

## Bar plot

- Load the `ggplot2::msleep` data set.

```{r}
ggplot2::msleep

```

- Draw a bar plot of number of observations (i.e., rows) for each taxonomic order
  (i.e, one plot and one bar per taxonomic order).

```{r}
ggplot(msleep, aes(x=order))+
  geom_bar()+
  theme(axis.text.x = element_text(size = 8, angle = 90, hjust = 1, vjust = 0.5))

```

- Change the angle and font size of the text for the x-axis ticks (not the axis titles).
  Justify the text of those x-axis ticks as right-aligned.
  (Copy-paste and extend the code chunk above.)

```{r}





```

- Change the value and font size of the title for both x and y axes.
  (Copy-paste and extend the code chunk above.)

```{r}

ggplot(msleep, aes(x=order))+
  geom_bar()+
  theme(axis.text.x = element_text(size = 8, angle = 90, hjust = 1, vjust = 0.5),
        axis.title = element_text(size = 12))






```

- Fill each bar with colors, proportionally to the count of each genus.
  (Copy-paste and extend the code chunk above.)

```{r}

ggplot(msleep, aes(x=order, fill = genus))+
  geom_bar()+
  theme(axis.text.x = element_text(size = 8, angle = 90, hjust = 1, vjust = 0.5),
        axis.title = element_text(size = 12))






```

From this point onwards, you may need to iteratively resize the text of the ticks and axes for readability.

- Reduce the legend key size.
  (Recommendation: use `unit(2, "mm")`).
  (Copy-paste and extend the code chunk above.)

```{r}
ggplot(msleep, aes(x=order, fill = genus))+
  geom_bar()+
  theme(axis.text.x = element_text(size = 8, angle = 90, hjust = 1, vjust = 0.5),
        axis.title = element_text(size = 12),
        legend.key.size = unit(2, "mm"))







```

- Force the legend to be display in 3 columns.
  (Recommendation: use `guide_legend(...)`).
  (Copy-paste and extend the code chunk above.)

```{r}
ggplot(msleep, aes(x=order, fill = genus))+
  geom_bar()+
  theme(axis.text.x = element_text(size = 8, angle = 90, hjust = 1, vjust = 0.5),
        axis.title = element_text(size = 12),
        legend.key.size = unit(2, "mm"))+
       guides(fill=guide_legend(ncol = 3))










```

- Add a contour of thin black lines to the bars.

```{r}

ggplot(msleep, aes(x=order, fill = genus))+
  geom_bar(colour="black", size = 0.2)+
  theme(axis.text.x = element_text(size = 8, angle = 90, hjust = 1, vjust = 0.5),
        axis.title = element_text(size = 12),
        legend.key.size = unit(2, "mm"))+
       guides(fill=guide_legend(ncol = 3))#as fill=genus is included in aes(), so use fill=guide_legend()








```

- Assign that last plot to an object called `obds_msleep`.

```{r}

obds_msleep <- ggplot(msleep, aes(x=order, fill = genus))+
  geom_bar(colour="black", size = 0.2)+
  theme(axis.text.x = element_text(size = 8, angle = 90, hjust = 1, vjust = 0.5),
        axis.title = element_text(size = 12),
        legend.key.size = unit(2, "mm"))+
       guides(fill=guide_legend(ncol = 2))









```

# Exercise

## Plotting grid

- Collate the plots that we assigend to objects through the day, as a single plot.

  + Plots: `obds_diamonds`, `obds_chickweight`, `obds_msleep`.
  
  + Methods: `cowplot::plot_grid()`, `patchwork`, `gridExtra::grid.arrange()`.

Using `cowplot`.

```{r, fig.height=7, fig.width=7}
library(cowplot)
first_row <- cowplot::plot_grid(obds_diamonds, obds_msleep, labels = c("A", "B"), ncol = 2, nrow = 1
)
print(first_row)
second_row <- cowplot::plot_grid(obds_chickweight, labels = "C")
print(second_row)
super_plot <- cowplot::plot_grid(first_row,second_row, nrow = 2
)
print(super_plot)

```

Using `patchwork`.

```{r, fig.height=12, fig.width=12}
library(patchwork)
super_plot_2 <- (obds_diamonds / obds_msleep) + obds_chickweight 

```

Using `gridExtra`.

```{r, fig.height=12, fig.width=12}
library(gridExtra)
super_plot_3 <- gridExtra::grid.arrange((obds_diamonds+theme_pubr()), (obds_chickweight+theme_pubr()), (obds_msleep+theme_pubr()+theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), legend.position = "right")), ncol = 2, nrow = 2)


```

- Export the new plot in a PDF file, and open it in a PDF viewer (e.g. Adobe Acrobat Reader DC).

You will likely need a few attempts to fine-tune the width and height of the output file.

```{r}
ggsave(
 "/project/obds/gliu/base_r/projects/ggplot2/super_plot.pdf", plot = super_plot, width = 14, height = 14 )
```
# load ggpubr package
```{r}
library(ggpubr)
```


# Exercise

## Pair programming

- Explore the data set `ggplot2::mpg` and generate the most informative plot that you can!

```{r}




```
